On Extending a Full-Sharing Design with Batched Scheduling

### 2015-04-15 paper.tex
implementation details -> end

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Implementations details}

At the implementation level, the major difference between local and
batched scheduling is in the tabling operation \emph{tabled new
  answer}, where we decide what to do when an answer is found during
the evaluation. This operation checks whether a newly found answer is
already in the corresponding answer trie structure and, if not,
inserts it. For the NS and SS designs the support for batched
scheduling is immediate, since the answer trie data structure is not
shared among threads. The usage of batched scheduling with the FS
design requires further support since with batched scheduling, answers
are immediately propagated and we have to ensure that the propagation
of an answer occurs on all subgoal calls one and only once. To do so,
we take advantage of the private chaining procedure, presented in the
previous subsection, as a way to keep, for every subgoal call of every
thread, track of all the answers that were already propagated. This
requires minor changes to the \emph{tabled new answer} tabling
operation. Algorithm~\ref{alg_table_new_answer_batched} shows how we
have extended the tabled new answer operation to support the FS design
with batched scheduling.

\begin{algorithm} [ht]
\caption{tabled\_new\_answer(answer ANS, subgoal frame SF)}
\algsetup{indent=0.65cm}
\begin{algorithmic}[1]
  \STATE $leaf \gets check\_insert\_answer\_trie(ANS, SF)$
  \IF {$NS\_design\ or\ SS\_design$}
    \STATE $. . .$                         \COMMENT{without changes}
  \ELSE                                                  [FS design]
    \STATE $chain \gets check\_insert\_consumer\_chain(leaf, SF)$
    \IF {$is\_answer\_marked\_as\_found(chain) = True$}
      \RETURN $failure$
     \ELSE                                        [the answer is new]
       \STATE $mark\_answer\_as\_found(chain)$
       \IF {$local\_scheduling\_mode(SF)$}
         \RETURN $failure$
       \ELSE                                    [batched scheduling mode]
         \RETURN $proceed$
       \ENDIF
     \ENDIF
   \ENDIF
\end{algorithmic}
\label{alg_table_new_answer_batched}
\end{algorithm}

The algorithm receives two arguments: the new answer found during the
evaluation (\emph{ANS}) and the subgoal frame which corresponds to the
call at hand (\emph{SF}). The \emph{NS\_design}, \emph{SS\_design} and
\emph{FS\_design} macros define which table design is enabled.

The algorithm begins by checking/inserting the given \emph{ANS} into
the answer trie structure, which will return the leaf node for the
path representing \emph{ANS} (line 1). In line 2, it then tests
whether one of the NS or SS designs are active, and in such a case,
the algorithm is remains unchanged.

Otherwise, for the FS design (lines 4 to 13), it checks/inserts
the given \emph{leaf} node into the private consumer chain for the
current thread, which will return the corresponding chain node. In
line 6, it then tests whether the chain node already existed in the
consumer chain, i.e., if it was inserted or not by the current
check/insert operation in order to return failure (line 7), or it
proceed with marking the answer \emph{ANS} has found (line 9). At the
end (lines 10 to 13), it returns failure if local scheduling is active
(line 11), otherwise, the batched scheduling is active, thus it
propagates the answer \emph{ANS} (line 13).

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Implementations Details}

At the implementation level, the major difference between local and
batched scheduling is in the tabling operation \emph{tabled new
  answer}, where we decide what to do when an answer is found during
the evaluation. This operation checks whether a newly found answer is
already in the corresponding answer trie structure and, if not,
inserts it. The usage of batched scheduling with the FS design
requires further support since with batched scheduling, answers are
immediately propagated and we have to ensure that the propagation of
an answer occurs on all subgoal calls one and only once. To do so, we
take advantage of the private chaining procedure, presented in the
previous subsection, as a way to keep, for every subgoal call of every
thread, track of all the answers that were already propagated. This
requires minor changes to the \emph{tabled new answer} tabling
operation. Algorithm~\ref{alg_table_new_answer_batched} shows how we
have extended the tabled new answer operation to support the FS design
with batched scheduling.

\begin{algorithm} [ht]
\caption{tabled\_new\_answer(answer ANS, subgoal frame SF)}
\algsetup{indent=0.65cm}
\begin{algorithmic}[1]
  \STATE $leaf \gets check\_insert\_answer\_trie(ANS, SF)$
  \STATE $chain \gets check\_insert\_consumer\_chain(leaf, SF)$
  \IF {$is\_answer\_marked\_as\_found(chain) = True$}
    \RETURN $failure$
  \ELSE                                        [the answer is new]
    \STATE $mark\_answer\_as\_found(chain)$
    \IF {$local\_scheduling\_mode(SF)$}
      \RETURN $failure$
    \ELSE                                [batched scheduling mode]
      \RETURN $proceed$
    \ENDIF
  \ENDIF  
\end{algorithmic}
\label{alg_table_new_answer_batched}
\end{algorithm}

The algorithm receives two arguments: the new answer found during the
evaluation (\emph{ANS}) and the subgoal frame which corresponds to the
call at hand (\emph{SF}). The algorithm begins by checking/inserting
the given \emph{ANS} into the answer trie structure, which will return
the leaf node for the path representing \emph{ANS} (line 1).

Then, it checks/inserts the given \emph{leaf} node into the private
consumer chain for the current thread, which will return the
corresponding chain node (line 2). Next in line 3, it tests whether
the chain node already existed in the consumer chain, i.e., if it was
inserted or not by the current check/insert operation in order to
return failure (line 4), or it proceeds with marking the answer
\emph{ANS} has found (line 6). At the end (lines 7 to 10), it returns
failure if local scheduling is active (line 8), otherwise, the batched
scheduling is active, thus it propagates the answer \emph{ANS} (line
10).
